---
title: "p8105_hw3_dz2399"
author: "DZ"
date: "October 4, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
```


## Problem 1
#### First, do some data cleaning:
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor” (i.e. no pre-collapsed categories)
* organize responses as a factor taking levels from “Excellent” to “Poor”
```{r}
data("brfss_smart2010")
df_brfss <- brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic== 'Overall Health') %>%
  select(-c(class,topic,question,sample_size,confidence_limit_low:geo_location)) %>% 
  mutate(response= as.factor(response))
df_brfss$response <-
  factor(df_brfss$response,levels(df_brfss$response)[c(4,2,3,5,1)])
rm(brfss_smart2010)
```

#### In 2002, which states were observed at 7 locations?
```{r}
df_brfss %>% 
  filter(year==2002) %>% 
  spread(key=response, value=data_value) %>% 
  group_by(locationabbr) %>% 
  mutate(locations = length(unique(locationdesc))) %>% 
  filter(locations == 7 ) %>% 
  distinct(locationabbr)
```

#### Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.
```{r}
df_brfss %>% 
  group_by(locationabbr,year) %>% 
  summarise(n_obs = length(data_value)) %>% 
  filter(year==2002|year==2010) %>% 
  ggplot(aes(x=year, y=n_obs, color = locationabbr))+
  geom_line()
  
  
```

#### Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r}
df_brfss %>% 
  spread(key=response, value=data_value) %>% 
  janitor::clean_names() %>% 
  filter(year==2002|year==2006|year==2010, locationabbr=='NY') %>% 
  group_by(year) %>% 
  summarise(mean=mean(excellent),
            sd=sd(excellent)) %>% 
  knitr::kable(digits = 1)
```

#### For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r}
df_brfss %>% 
  group_by(year, locationabbr, response) %>% 
  summarise(state_mean=mean(data_value)) %>% 
  ggplot(aes(x=year,y=state_mean))+
  geom_point()+
  facet_grid(.~response)
```

## Problem 2

#### The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

```{r, include=F}
data("instacart")
```

The dataset has the dimension of `r dim(instacart)`. The order_id and product_is is a numeric variable that is the order identifier and the product identifier. The add_to_cart_order is a numeric value that represents the order in which each product was added to cart. 

#### How many aisles are there, and which aisles are the most items ordered from?
```{r}

instacart <- instacart %>% 
  janitor::clean_names()

instacart %>% 
  distinct(aisle_id) %>% 
  nrow()

instacart %>% 
  group_by(aisle) %>% 
  summarise(number_of_items = n()) %>% 
  arrange(desc(number_of_items)) %>% 
  slice(1)
```

#### Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarise(number_of_items = n()) %>% 
  arrange(desc(number_of_items)) %>% 
  ggplot(aes(x=aisle,y=number_of_items,size=number_of_items))+geom_point()
```

#### Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
```{r}
instacart %>% 
  filter(aisle=='baking ingredients'|aisle== 'dog food care'|aisle== 'packaged vegetables fruits') %>% 
  group_by(aisle,product_name) %>% 
  summarise(num_of_purchase = n()) %>% 
  group_by(aisle) %>% 
  filter(num_of_purchase == max(num_of_purchase)) %>% 
  knitr::kable()
```

#### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
```{r}
instacart %>% 
  filter(product_name=='Pink Lady Apples'
         |product_name=='Coffee Ice Cream') %>% 
  select(order_dow,order_hour_of_day,product_name) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_time= mean(order_hour_of_day)) %>% 
  spread(key=product_name, value=mean_time)
```

## Problem 3
```{r,include=F}
data("ny_noaa")
```


#### Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
df_nynoaa <- ny_noaa %>% 
  separate(date,into = c("year","month","day")) %>% 
  janitor::clean_names() %>% 
  mutate(tmin=as.numeric(tmin),
         tmax=as.numeric(tmax)) %>% 
  na.omit()
  
df_nynoaa %>% 
  group_by(snow) %>% 
  summarise(obs=n()) %>% 
  filter(obs==max(obs))
```


#### Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
df_nynoaa %>% 
  filter(month=='01'|month=='06') %>%
  mutate(tmean = (tmax+tmin)/2) %>% 
  group_by(month,year) %>% 
  summarise(month_mean=mean(tmean)) %>% 
  ggplot(aes(x=year,y=month_mean))+
  geom_point()+
  facet_grid(.~month,scales = 'free')
  
```

